---
title: "Green Jobs and Growth"
subtitle: "Making the West of England the home for green jobs and growth"
author: "SC"
date: last-modified
execute:
  echo: false
  warning: false
  message: false
---

## Context

This indicator summary tracks indicators related to green economy and environmental sustainability across the West of England. Topics include carbon emissions, renewable energy, water pollution, climate resilience and public health. Indicators were developed to provide high-level insights into regional trends and performance in relation to the West of England's environmental priorities, which are informed by the [Growth Strategy](https://www.westofengland-ca.gov.uk/wp-content/uploads/2025/09/WE4915-Growth-Strategy-exec-summary_aw_web_v1.pdf).

```{r}
#| label: setup
#| code-summary: "Load required packages and source files"
#| include: false

# Load required packages
library(tidyverse)

# Source shared helpers (theme, helpers, fact_helpers, collate_fact)
source(here::here("scripts", "R", "_common.R"))
source(here::here("scripts", "R", "05-environment", "RI_5_ghg_emissions.R"))
source(here::here("scripts", "R", "05-environment", "RI_5A1_renewable.R"))
source(here::here("scripts", "R", "05-environment", "RI_5B1_nondom_epc_a.R"))
source(here::here(
  "scripts",
  "R",
  "05-environment",
  "RI_5E1_heat_stress_deaths.R"
))
source(here::here(
  "scripts",
  "R",
  "05-environment",
  "RI_5E2_flood_warning_index.R"
))
source(here::here("scripts", "R", "05-environment", "RI_5F1_tree_canopy.R"))
source(here::here(
  "scripts",
  "R",
  "05-environment",
  "RI_5F4_water_catchment_health.R"
))

indicator_fact_tbl <- collate_fact() |>
  build_reporting_view()

indicator_polarity <- core_dim_data_tbl |>
  dplyr::select(indicator_id, polarity) |>
  tibble::deframe()
```

```{r}
#| label: priority_summary
#| eval: true
#| code-summary: "Priority 5 indicator summary table"
format_priority_summary(
  indicator_fact_tbl,
  core_dim_data_tbl,
  priority = 5,
  title = "Green Jobs and Growth: Indicator Summary"
)
```
## Greenhouse Gas Emissions

```{r}
#| label: RI_5_ghg_emissions_gt
#| eval: true
#| code-summary: "RI_5_ghg_emissions KPI summary"
indicator_fact_tbl |>
  format_indicator_summary(
    "RI_5_ghg_emissions",
    dim_tbl = core_dim_data_tbl,
    title = "Greenhouse Gas Emissions",
    subtitle = "Transport, Commercial, Industry, Public sector, Domestic",
    polarity = indicator_polarity[["RI_5_ghg_emissions"]]
  )
```

### Overview

Regional greenhouse gas emissions from transport, domestic, commercial, industrial and public sectors. This tracks the region's progress toward the net zero target by 2030 and identifies key sectors for emissions reduction. Source data are [UK local authority and regional greenhouse gas emissions statistics, 2005 to 2023](https://www.gov.uk/government/collections/uk-local-authority-and-regional-greenhouse-gas-emissions-statistics) from the Department for Energy Security and Net Zero.

```{r}
#| label: RI_5_ghg_emissions
#| eval: true
#| code-summary: "RI_5_ghg_emissions analysis"
#| fig-alt: |
#|   Stacked bar chart of greenhouse gas emissions by sector in the West of
#|   England Combined Authority area, 2014 to 2023. The y-axis shows emissions
#|   in kilotonnes of CO2 equivalent; the x-axis shows year. Five sectors are
#|   colour-coded: Transport (green, largest), Domestic (red), Commercial
#|   (dark green), Industry (purple), and Public Sector (black). Total
#|   emissions declined from around 6,000 kt CO2e in 2014 to approximately
#|   4,400 kt CO2e in 2023, though the region is not on track to meet its 2030
#|   net zero target. Transport and domestic heating remain the dominant
#|   sectors.
RI_5_plot
```
The indicator is defined by subset of sectors. To calculate the proportion of total emissions contributed by each sector, the current year total emissions are calculated and then used to derive percentage contributions for the current year. This allows for alignment with other work in the combined authority supporting transport and domestic sector decarbonisation. Note that LULUCF (land use, land use change and forestry) emissions are negative due to carbon sequestration and percentages are calculated based on the final balance after removals are subtracted.

```{r}
#| label: RI_5_ghg_emissions_sector_proportions_current_year
#| eval: true
#| code-summary: "RI_5_ghg_emissions sector proportions current year"
#| fig-alt: |
#|   Bar chart showing the percentage contribution of each sector to total
#|   greenhouse gas emissions in the West of England Combined Authority area for
#|   the latest year - typically 2 years behind the current year. The y-axis shows sector; the x-axis shows
#|   emissions in KT CO2e. The chart shows that Transport is the largest contributor followed by Domestic
RI_5_sector_current_year_plot
```
## Heat Associated Mortality

```{r}
#| label: RI_5E1_heat_stress_deaths_gt
#| eval: true
#| code-summary: "summary"

indicator_fact_tbl |>
  format_indicator_summary(
    "RI_5E1_heat_stress_deaths",
    dim_tbl = core_dim_data_tbl,
    title = "Heat Associated Mortality",
    subtitle = "Deaths per million population",
    polarity = indicator_polarity[["RI_5E1_heat_stress_deaths"]]
  )
```

### Overview
This indicator tracks the number of heat associated deaths per million population in the Avon Local Resilience Forum area, which largely correlates with the West of England. It provides insight into the region's vulnerability to heat stress through the lens of public health. Data for the region are only available since 2024.

```{r}
#| label: RI_5E1_heat_stress_deaths_plot
#| eval: true
#| code-summary: "RI_5E1_heat_stress_deaths plot"
#| fig-alt: |
#|   Line and point chart showing heat-associated mortality in the Avon Local
#|   Resilience Forum area for 2024 and 2025. The y-axis shows deaths per
#|   million population on a scale from 0 to 50; the x-axis shows year. Two
#|   data points are shown: approximately 33 deaths per million in 2024 rising
#|   to around 35 in 2025. The small increase is consistent with the
#|   continuing rise in ambient temperatures from a warming climate.
RI_5E1_heat_stress_deaths_plot
```
## Flood Warnings Intensity

```{r}
#| label: RI_5E2_flood_warning_index_gt
#| eval: true
#| code-summary: "summary"

indicator_fact_tbl |>
  format_indicator_summary(
    "RI_5E2_flood_warning_index",
    dim_tbl = core_dim_data_tbl,
    title = "Flood Warnings Intensity Index",
    subtitle = "Relative to 2020 baseline",
    polarity = indicator_polarity[["RI_5E2_flood_warning_index"]]
  )
```

### Overview

This indicator is designed to represent the risk of flooding relative to a 2020 baseline. It is a duration-weighted composite indicator tracking flood warning activity across the West of England (Bristol, Bath & North East Somerset, South Gloucestershire, North Somerset and derived from the Environment Agency's [historic flood warnings dataset](https://environment.data.gov.uk/dataset/88bed270-d465-11e4-8669-f0def148f590). It provides annual monitoring to support climate resilience reporting. The methodology for the indicator is described in this [github repository](https://github.com/stevecrawshaw/fwii).

```{r}
#| label: RI_5E2_flood_warning_index_plot
#| eval: true
#| code-summary: "RI_5E2_flood_warning_index plot"
#| fig-alt: |
#|   Line and point chart of the Flood Warning Intensity Index (FWII) for the
#|   West of England from 2020 to 2025. The y-axis shows the composite index
#|   value on a scale from 0 to 200; the x-axis shows year. A dashed green
#|   horizontal line marks the 2020 baseline at 100. The index fell sharply to
#|   around 40 in 2021 and 30 in 2022, returned to the baseline in 2023,
#|   peaked at approximately 165 in 2024, then dropped below the baseline to
#|   around 70 in 2025. The 2024 peak represents the highest flood warning
#|   intensity in the series.
RI_5E2_flood_warning_index_plot
```
## Water Catchment Health

```{r}
#| label: RI_5F4_water_catchment_health_gt
#| eval: true
#| code-summary: "RI_5F4_water_catchment_health summary"

indicator_fact_tbl |>
  format_indicator_summary(
    "RI_5F4_water_catchment_health",
    dim_tbl = core_dim_data_tbl,
    title = "Water Catchment Health",
    subtitle = "Proportion of surface waters in 'Good' ecological health",
    polarity = indicator_polarity[["RI_5F4_water_catchment_health"]]
  )
```

### Overview

This indicator tracks the ecological health of surface waters in the catchment which most closely represents the region - [Avon Bristol and Somerset North Streams](https://environment.data.gov.uk/catchment-planning/ManagementCatchment/3005/classifications). It is expressed as the proportion of surface waters in 'Good' ecological health. The data are derived from the Environment Agency's regular water quality monitoring programme. Surface water classification and sources have changed over the years, so data prior to 2019 are not comparable with recent datasets.

```{r}
#| label: RI_5F4_water_catchment_health_plot
#| eval: true
#| code-summary: "RI_5F4_water_catchment_health plot"
#| fig-alt: |
#|   Line and point chart of the proportion of surface waters in 'Good'
#|   ecological health in the Avon Bristol and Somerset North Streams
#|   catchment, for 2019 and 2022. The y-axis shows percentage from 0% to
#|   100%; the x-axis shows year. Both data points sit at approximately 14-15%,
#|   showing virtually no change and reflecting the national picture of
#|   persistently poor surface water quality.

RI_5F4_plot
```
## Tree Canopy

```{r}
#| label: RI_5F1_tree_canopy_gt
#| eval: true
#| code-summary: "summary"

indicator_fact_tbl |>
  format_indicator_summary(
    "RI_5F1_tree_canopy",
    dim_tbl = core_dim_data_tbl,
    title = "Tree Canopy Cover",
    subtitle = "National Forest Inventory Woodland",
    polarity = indicator_polarity[["RI_5F1_tree_canopy"]]
  )
```

### Overview

This indicator tracks the extent of tree canopy cover in the region, as measured by the National Forest Inventory (NFI). Tree canopy cover is an important aspect of urban greening, biodiversity and climate resilience. The NFI source [dataset](https://data-forestry.opendata.arcgis.com/datasets/d62fd97664a64596844b9834c1737e46_0/explore) is filtered by the "Woodland" category and clipped to the boundary of the West of England to provide a regional measure of the area covered by woodland.

```{r}
#| label: RI_5F1_tree_canopy_plot
#| eval: true
#| code-summary: "RI_5F1_tree_canopy plot"
#| fig-alt: |
#|   Line chart showing the percentage change in National Forest Inventory
#|   woodland extent in the West of England from the 2020 baseline, covering
#|   2020 to 2024. The y-axis shows percentage change on a scale from -1% to
#|   +1%; the x-axis shows year. The line remains very close to zero
#|   throughout, ending at approximately +0.3% by 2024, indicating a small
#|   but positive increase in woodland cover since 2020.

RI_5F1_plot
```
## Non-Domestic Energy Performance

```{r}
#| label: RI_5B1_nondom_epc_a_gt
#| eval: true
#| code-summary: "summary"

indicator_fact_tbl |>
  format_indicator_summary(
    "RI_5B1_nondom_epc_a",
    dim_tbl = core_dim_data_tbl,
    title = "Non - Domestic Energy Ratings",
    subtitle = "Proportion of properties rated A or A+",
    polarity = indicator_polarity[["RI_5B1_nondom_epc_a"]]
  )
```

### Overview

This indicator measures the energy efficiency of non - domestic property by expressing the proportion of properties which have an energy rating of A or A+, in other words the best performing properties. Data are sourced from the [Ministry of Housing, Communities and Local Government](https://get-energy-performance-data.communities.gov.uk/).

```{r}
#| label: RI_5B1_nondom_epc_a_plot
#| eval: true
#| code-summary: "RI_5B1_nondom_epc_a monthly cumulative plot"
#| fig-alt: |
#|   Line chart showing the cumulative monthly proportion of non-domestic
#|   properties in the West of England with an EPC energy rating of A or A+,
#|   from 2017 to 2026. The y-axis shows percentage from 0% to 100%; the
#|   x-axis shows date. The line rises very gradually from under 2% to
#|   approximately 6% by 2026, showing slow but steady improvement in the
#|   proportion of best-performing non-domestic buildings.
RI_5B1_nondom_epc_a_plot
```
## Domestic Renewable Energy

```{r}
#| label: RI_5A2_domestic_renewable_gt
#| eval: true
#| code-summary: "RI_5A2_domestic_renewable summary"
# no specific R script for this analysis - it's in RI_5A1
indicator_fact_tbl |>
  format_indicator_summary(
    "RI_5A2_domestic_renewable",
    dim_tbl = core_dim_data_tbl,
    title = "Domestic Renewable Energy",
    subtitle = "Photovoltaic sites",
    polarity = indicator_polarity[["RI_5A2_domestic_renewable"]]
  )
```
### Overview

Solar photovoltaic (PV) installations are a key aspect of energy security and net zero. This indicator tracks installations per household in the region. Almost all domestic PV installations are rooftop, so the indicator is expressed in terms of site counts rather than capacity. The data are derived from the Department for Energy Security and Net Zero's [renewable energy by local authority spreadsheet](https://assets.publishing.service.gov.uk/media/68da76d2c487360cc70c9e9d/Renewable_electricity_by_local_authority_2014_-_2024.xlsx).

```{r}
#| label: RI_5A2_domestic_renewable_plot
#| eval: true
#| code-summary: "RI_5A2_domestic_renewable plot bar chart"
#| fig-alt: |
#|   Stacked bar chart of photovoltaic (PV) site counts by local authority in
#|   the West of England, from 2015 to 2024. The y-axis shows the number of
#|   sites; the x-axis shows year. North Somerset (black) has the largest
#|   share, followed by Bristol (red), Bath and North East Somerset (purple),
#|   and South Gloucestershire (dark green). Total PV sites increased steadily
#|   from around 16,000 in 2015 to over 36,000 in 2024, indicating consistent
#|   growth in domestic solar adoption.
RI_5A2_domestic_renewable_plot
```
## Renewable Energy

```{r}
#| label: RI_5A1_renewable_gt
#| eval: true
#| code-summary: "RI_5A1_renewable summary"

indicator_fact_tbl |>
  format_indicator_summary(
    "RI_5A1_renewable_capacity_km2",
    dim_tbl = core_dim_data_tbl,
    title = "Renewable Electricity Installed Capacity",
    subtitle = "Installed capacity per square Km",
    polarity = indicator_polarity[["RI_5A1_renewable_capacity_km2"]]
  )
```
### Overview

Potential for renewable energy capacity is strongly influenced by land availability and price, hence this indicator is expressed in terms of Megawatts per square Kilometre to enable future comparison with other regions.

```{r}
#| label: RI_5A1_renewable_plot
#| eval: true
#| code-summary: "RI_5A1_renewable plot bar chart"
#| fig-alt: |
#|   Stacked bar chart of renewable electricity installed capacity in
#|   Megawatts by local authority in the West of England, from 2015 to 2024.
#|   The y-axis shows capacity in MW; the x-axis shows year. South
#|   Gloucestershire (dark green) contributes the largest share, followed by
#|   North Somerset (black), Bristol (red), and Bath and North East Somerset
#|   (purple). Total installed capacity grew steadily from around 250 MW in
#|   2015 to over 525 MW in 2024, reflecting the region's ambitions for net
#|   zero.

RI_5A1_renewable_plot
```
