---
title: "Economic Growth"
subtitle: "Contributing to national economic growth helping our businesses succeed and creating jobs"
author: "SN"
date: last-modified
---

## Context

This indicator summary tracks regional indicators related to the West of England's contribution to national economic growth. It includes indicators related to business growth, productivity, and employment. It draws on a range of national and local data sources to provide a high-level overview of regional trends and performance in relation to economic growth. Indicators were defined as part of the development of our [Growth Strategy](https://www.westofengland-ca.gov.uk/wp-content/uploads/2025/09/WE4915-Growth-Strategy-exec-summary_aw_web_v1.pdf) and are intended to provide a high-level overview of regional trends and performance in relation to economic growth.

```{r}
#| label: 01_setup
#| code-summary: "Load required packages and source files"
#| include: false

# Load required packages
library(tidyverse)

# Source shared helpers (theme, helpers, fact_helpers, collate_fact)
source(here::here("scripts", "R", "_common.R"))
source(here::here(
  "scripts",
  "R",
  "01-economy",
  "RI_1A2_employment_priority_sectors.R"
))
source(here::here("scripts", "R", "01-economy", "RI_1B1_gdp_growth.R"))
source(here::here("scripts", "R", "01-economy", "RI_1B2_gva.R"))
source(here::here("scripts", "R", "01-economy", "RI_1A1_employment_rate.R"))
source(here::here("scripts", "R", "01-economy", "RI_1B3_median_pay.R"))
source(here::here("scripts", "R", "01-economy", "RI_1D1_broadband_coverage.R"))
source(here::here("scripts", "R", "01-economy", "RI_1E2_population_change.R"))
source(here::here("scripts", "R", "01-economy", "RI_1E3_area_satisfaction.R"))

indicator_fact_tbl <- collate_fact() |>
  build_reporting_view()

indicator_polarity <- core_dim_data_tbl |>
  dplyr::select(indicator_id, polarity) |>
  tibble::deframe()
```
```{r}
#| label: priority_summary
#| eval: true
#| code-summary: "Priority 1 indicator summary table"
format_priority_summary(
  indicator_fact_tbl,
  core_dim_data_tbl,
  priority = 1,
  title = "Economic Growth: Indicator Summary"
)
```
## Real GDP growth rate

```{r}
#| label: RI_1B1_real_gdp_growth
#| eval: true
#| code-summary: "RI_1B1_real_gdp_growth KPI summary"
indicator_fact_tbl |>
  format_indicator_summary(
    "RI_1B1_gdp_growth",
    dim_tbl = core_dim_data_tbl,
    title = "GDP growth since 2014",
    subtitle = "Real GDP (excluding inflation) for the West of England",
    polarity = indicator_polarity[["RI_1B1_gdp_growth"]]
  )
```

### Overview

Growth in Gross Domestic Product (GDP) since 2014 based in 2023 prices (in essence, excluding inflation). This tracks the region's progress in relation to its aim of increasing GDP by 28% over 10 years.

```{r}
#| label: RI_1B1_real_gdp_growth_plot
#| eval: true
#| code-summary: " RI_1B1_real_gdp_growth analysis"
#| fig-alt: |
#|   Line chart of a real GDP index for the West of England, 2014 to 2023
#|   The line rises in most years, showing sustained
#|   growth in real GDP over the decade as the region tracks its target of
#|   a 28% increase.
RI_1B1_plot
```

## GVA per job

```{r}
#| label: RI_1B2_gva_per_job
#| eval: true
#| code-summary: "RI_1B2_gva_per_job KPI summary"
indicator_fact_tbl |>
  format_indicator_summary(
    "RI_1B2_gva",
    dim_tbl = core_dim_data_tbl,
    title = "Annual GVA per job, 2013-2023",
    subtitle = "Current price GVA (including inflation) per job for the West of England",
    polarity = indicator_polarity[["RI_1B2_gva"]]
  )
```

### Overview

Current price GVA per job is a measure of productivity. It provides insights into how much a worker can produce in their job and it is a key measure of economic growth.

```{r}
#| label: RI_1B2_gva_per_job_plot
#| eval: true
#| code-summary: "RI_1B2_gva_per_job analysis"
#| fig-alt: |
#|   Line chart of annual current-price GVA per job in pounds for the West
#|   of England, 2013 to 2023. The line trends upward over the period,
#|   indicating rising productivity per worker across the region.
RI_1B2_plot
```

## Employment Rate 16-64

```{r}
#| label: RI_1A1_employment_rate_gt
#| eval: true
#| code-summary: "RI_1A1_employment_rate summary"

indicator_fact_tbl |>
  format_indicator_summary(
    "RI_1A1_employment_rate",
    dim_tbl = core_dim_data_tbl,
    title = "Employment Rate 16-64",
    subtitle = "Proportion of working-age residents in employment",
    polarity = indicator_polarity[["RI_1A1_employment_rate"]]
  )
```

### Overview

Employment is a key measure of economic participation and prosperity. This indicator measures the proportion of working-age residents who are in employment. Higher employment rates indicate that residents are benefiting from economic opportunities and contributing their skills within the labour market. Sustaining high employment levels is important for supporting inclusive growth and regional productivity.

```{r}
#| label: RI_1A1_employment_rate_plot
#| eval: true
#| code-summary: "RI_1A1_employment_rate line chart"
#| fig-alt: |
#|   Line chart of the employment rate for residents aged 16 to 64, shown
#|   as two lines: the West of England in green and Great Britain in grey,
#|   covering the latest available years. The two lines track closely
#|   together, with the West of England generally at or slightly above the
#|   Great Britain rate.

RI_1A1_employment_rate_plot
```

## Employment level and growth for priority sectors

```{r}
#| label: RI_1A2_employment_priority_sectors
#| eval: true
#| code-summary: "RI_1A2_employment_priority_sectors KPI summary"
indicator_fact_tbl |>
  format_indicator_summary(
    "RI_1A2_employment_priority_sectors",
    dim_tbl = core_dim_data_tbl,
    title = "Priority sector employment",
    subtitle = "West of England",
    polarity = indicator_polarity[["RI_1A2_employment_priority_sectors"]]
  )
```

### Overview

Total employment in priority sectors, excluding clean energy. This tracks the region's progress in relation to its aim of increasing employment in priority sectors by 72,000 jobs over 10 years. Clean energy is excluded from this total as the priority sectors are measured using Standard Industry Classification codes (SICs) and there are no SICs for clean energy.


```{r}
#| label: RI_1A2_employment_priority_sectors_plot
#| eval: true
#| code-summary: "RI_1A2_employment_priority_sectors analysis"
#| fig-alt: |
#|   Line chart of total employment in priority sectors, excluding clean
#|   energy, for the West of England from 2015 to the most recent year.
#|   The line shows the region's progress towards its target of adding
#|   72,000 priority sector jobs over 10 years.
RI_1A2_employment_priority_sectors_plot
```

## Gross monthly median wage

```{r}
#| label: RI_1B3_median_wage
#| eval: true
#| code-summary: " RI_1B3_median_wage KPI summary"
indicator_fact_tbl |>
  format_indicator_summary(
    "RI_1B3_median_pay",
    dim_tbl = core_dim_data_tbl,
    title = "Annual GVA per job, 2013-2023",
    subtitle = "Current price GVA (including inflation) per job for the West of England",
    polarity = indicator_polarity[["RI_1B3_median_pay"]]
  )
```

### Overview

Gross monthly median wage provides an indication of how wages are changing across the region. The median relates to the person in the middle of the distribution, so it is more reliable than a mean average, which can be skewed upwards by those on very high wages. Whilst a rising median wage is a good thing, the measure takes no account of how wages are changing across the distribution. For example, those on very low wages might see their pay change at a different rate to those in other pay brackets, but this is not reflected in the median.

```{r}
#| label: RI_1B3_median_wage_plot
#| eval: true
#| code-summary: "RI_1B3_median_wage analysis"
#| fig-alt: |
#|   Line chart of the estimated median monthly wage in pounds for the West
#|   of England over the most recent 10 years. The line rises over time,
#|   showing wages increasing across the region, though this measure does
#|   not show how the increase is distributed between low and high earners.
RI_1B3_plot
```

## Broadband coverage

```{r}
#| label: RI_1D1_broadband_coverage
#| eval: true
#| code-summary: "RI_1D1_broadband_coverage KPI summary"
indicator_fact_tbl |>
  format_indicator_summary(
    "RI_1D1_broadband_coverage",
    dim_tbl = core_dim_data_tbl,
    title = "Superfast coverage",
    subtitle = "Residential and commercial premises",
    polarity = indicator_polarity[["RI_1D1_broadband_coverage"]]
  )
```

### Overview

Regional broadband coverage for residential and commercial premises. This tracks the region's progress in relation to superfast broadband coverage (30mbps+) and gigabit broadband coverage (1000mbps+).

```{r}
#| label: RI_1D1_broadband_coverage_plot
#| eval: true
#| code-summary: "RI_1D1_broadband_coverage analysis"
#| fig-alt: |
#|   Line chart showing two coverage measures for the West of England over
#|   time: superfast broadband (30mbps+) in green and gigabit broadband
#|   (1000mbps+) in purple. Superfast coverage is consistently higher and
#|   close to full coverage, while gigabit coverage started lower but has
#|   been catching up in more recent years.
RI_1D1_plot
```

## Population change

```{r}
#| label: RI_1E2_population_change
#| eval: true
#| code-summary: "RI_1E2_population_change KPI summary"
indicator_fact_tbl |>
  format_indicator_summary(
    "RI_1E2_population_change",
    dim_tbl = core_dim_data_tbl,
    title = "Total population change (cumulative)",
    subtitle = "West of England",
    polarity = indicator_polarity[["RI_1E2_population_change"]]
  )
```

### Overview

Cumulative population change between 2016 and 2025 in the West of England, including constituent authorities and North Somerset. This tracks the rate of change in the region's total population.

```{r}
#| label: RI_1E2_population_change_plot
#| eval: true
#| code-summary: "RI_1E2_population_change analysis"
#| fig-alt: |
#|   Line chart of cumulative percentage population change from 2016 to
#|   2025, with separate coloured lines for Bath and North East Somerset,
#|   Bristol, North Somerset, South Gloucestershire, and the West of
#|   England total. All areas show a rising cumulative trend, with the
#|   individual authorities spread above and below the regional total.
RI_1E2_plot
```

## Satisfaction with local area as a place to live

```{r}
#| label: RI_1E3_area_satisfaction
#| eval: true
#| code-summary: "RI_1E3_area_satisfaction KPI summary"
indicator_fact_tbl |>
  format_indicator_summary(
    "RI_1E3_area_satisfaction",
    dim_tbl = core_dim_data_tbl,
    title = "Satisfaction with local area as a place to live",
    subtitle = "Bristol only - see plot for all constituent authorities and North Somerset",
    polarity = indicator_polarity[["RI_1E3_area_satisfaction"]]
  )
```

### Overview

Respondent satisfaction with their local area as a place to live as reported by the Community Life Survey. These results are published by the Department for Culture, Media and Sport. These data track respondent satisfaction with their local area as a place to live at unitary authority level.

```{r}
#| label: RI_1E3_area_satisfaction_plot
#| eval: true
#| code-summary: "RI_1E3_area_satisfaction analysis"
#| fig-alt: |
#|   Line chart of the percentage of respondents satisfied with their local
#|   area as a place to live, with separate coloured lines for Bath and
#|   North East Somerset, Bristol, North Somerset, and South Gloucestershire
#|   over time. Satisfaction levels are broadly high across all four
#|   authorities, with some year-to-year variation between them.
RI_1E3_plot
```


