---
title: "Economic Growth"
subtitle: "Contributing to national economic growth helping our businesses succeed and creating jobs"
author: "SN"
date: last-modified
execute:
  echo: false
  warning: false
  message: false
---

## Context

This chapter tracks regional indicators related to the West of England's contribution to national economic growth. It includes indicators related to business growth, productivity, and innovation. The chapter draws on a range of national and local data sources to provide a high-level overview of regional trends and performance in relation to economic growth. Indicators were defined as part of the development of our [Growth Strategy](https://www.westofengland-ca.gov.uk/wp-content/uploads/2025/09/WE4915-Growth-Strategy-exec-summary_aw_web_v1.pdf) and are intended to provide a high-level overview of regional trends and performance in relation to economic growth.

```{r}
#| label: 01_setup
#| code-summary: "Load required packages and source files"
#| include: false

# Load required packages
library(tidyverse)

# Source shared helpers (theme, helpers, fact_helpers, collate_fact)
source(here::here("scripts", "R", "_common.R"))
# source(here::here("scripts", "R", "load_pins.R"))
source(here::here("scripts", "R", "01-economy", "RI_1D1_broadband_coverage.R"))
source(here::here("scripts", "R", "01-economy", "RI_1E2_population_change.R"))
source(here::here("scripts", "R", "01-economy", "RI_1E3_area_satisfaction.R"))

indicator_fact_tbl <- collate_fact() |>
  build_reporting_view()

indicator_polarity <- core_dim_data_tbl |>
  dplyr::select(indicator_id, polarity) |>
  tibble::deframe()
```

---

## Broadband coverage

```{r}
#| label: RI_1D1_broadband_coverage
#| eval: true
#| code-summary: "RI_1D1_broadband_coverage KPI summary"
indicator_fact_tbl |>
  format_indicator_summary(
    "RI_1D1_broadband_coverage",
    title = "Superfast coverage",
    subtitle = "Residential and commercial premises",
    units = "percent",
    polarity = indicator_polarity[["RI_1D1_broadband_coverage"]]
  )
```

### Overview

Regional broadband coverage for residential and commercial premises. This tracks the region's progress in relation to superfast broadband coverage (30mbps+) and gigabit broadband coverage (1000mbps+).

```{r}
#| label: RI_1D1_broadband_coverage_plot
#| eval: true
#| code-summary: "RI_1D1_broadband_coverage analysis"
RI_1D1_plot
```

---

---

## Population change

```{r}
#| label: RI_1E2_population_change
#| eval: true
#| code-summary: "RI_1E2_population_change KPI summary"
indicator_fact_tbl |>
format_indicator_summary(
  "RI_1E2_population_change",
  title = "Total population change (cumulative)",
  subtitle = "West of England",
  units = "percent",
  polarity = indicator_polarity[["RI_1E2_population_change"]]
)
```

### Overview

Cumulative population change between 2016 and 2025 in the West of England, including constituent authorities and North Somerset. This tracks the rate of change in the region's total population.

```{r}
#| label: RI_1E2_population_change_plot
#| eval: true
#| code-summary: "RI_1E2_population_change analysis"
RI_1E2_plot
```

---

---

## Satisfaction with local area as a place to live

```{r}
#| label: RI_1E3_area_satisfaction
#| eval: true
#| code-summary: "RI_1E3_area_satisfaction KPI summary"
indicator_fact_tbl |>
format_indicator_summary(
  "RI_1E3_area_satisfaction",
  title = "Satisfaction with local area as a place to live",
  subtitle = "Bristol only - see plot for all constituent authorities and North Somerset",
  units = "percent",
  polarity = indicator_polarity[["RI_1E3_area_satisfaction"]]
)
```

### Overview

Respondent satisfaction with their local area as a place to live as reported by the Community Life Survey. These results are published by the Department for Culture, Media and Sport. These data track respondent satisfaction with their local area as a place to live at unitary authority level.

```{r}
#| label: RI_1E3_area_satisfaction_plot
#| eval: true
#| code-summary: "RI_1E3_area_satisfaction analysis"
RI_1E3_plot
```

---
