---
title: "Future-Ready Skills"
subtitle: "Empowering residents with future-ready skills"
author: "MJ"
date: last-modified
execute:
  echo: false
  warning: false
  message: false
---

## Context

This indicator summary tracks indicators related to skills development and training across the West of England. Indicators cover metrics like qualification levels, apprenticeships, adult learning participation, skills gaps, and employment outcomes. Indicators were developed to provide high-level insights into regional trends and performance in relation to the West of England's skills priorities, which are informed by the [Growth Strategy](https://www.westofengland-ca.gov.uk/wp-content/uploads/2025/09/WE4915-Growth-Strategy-exec-summary_aw_web_v1.pdf).

```{r}
#| label: setup
#| code-summary: "Load required packages and source files"
#| include: false

# Load required packages
library(tidyverse)

# Source shared helpers (theme, helpers, fact_helpers, collate_fact)
source(here::here("scripts", "R", "_common.R"))

#Sourcing indicators
source(here::here("scripts", "R", "04-skills", "RI_4A1_no_qualifications.R"))
source(here::here("scripts", "R", "04-skills", "RI_4A2_level4_plus.R"))
source(here::here(
  "scripts",
  "R",
  "04-skills",
  "RI_4A3_apprenticeship_starts.R"
))
source(here::here("scripts", "R", "04-skills", "RI_4A4_ks4_attainment8.R"))
# source(here::here(
#   "scripts",
#   "R",
#   "04-skills",
#   "RI_4B1_employment_rate_16_64.R"
# ))
source(here::here("scripts", "R", "04-skills", "RI_4B2_economic_inactivity.R"))
source(here::here("scripts", "R", "04-skills", "RI_4B3_neet_16_17.R"))
source(here::here("scripts", "R", "04-skills", "RI_4B4_neet_18_24.R"))
source(here::here("scripts", "R", "04-skills", "RI_4B5_below_living_wage.R"))
source(here::here(
  "scripts",
  "R",
  "04-skills",
  "RI_4B6_claimant_deprived_wards.R"
))
source(here::here("scripts", "R", "04-skills", "RI_4C1_skills_gaps.R"))

indicator_fact_tbl <- collate_fact() |>
  build_reporting_view()

indicator_polarity <- core_dim_data_tbl |>
  dplyr::select(indicator_id, polarity) |>
  tibble::deframe()
```

```{r}
#| label: priority_summary
#| eval: true
#| code-summary: Priority 4 indicator summary table
format_priority_summary(
  indicator_fact_tbl,
  core_dim_data_tbl,
  priority = 4,
  title = "Future-Ready Skills: Indicator Summary"
)
```

## Residents with No Qualifications

```{r}
#| label: RI_4A1_no_qualifications_gt
#| eval: true
#| code-summary: "RI_4A1_no_qualifications summary"

indicator_fact_tbl |>
  format_indicator_summary(
    "RI_4A1_no_qualifications",
    dim_tbl = core_dim_data_tbl,
    title = "Residents with No Qualifications",
    subtitle = "Working-age population with no formal qualifications",
    polarity = indicator_polarity[["RI_4A1_no_qualifications"]]
  )
```

### Overview

Qualifications are a key driver of employment, productivity and economic opportunity. This indicator measures the proportion of working-age residents with no formal qualifications. Lower levels indicate a more skilled workforce and greater access to employment opportunities. Reducing the number of residents without qualifications supports the West of England Growth Strategy's ambition to equip people with the skills needed to access future jobs, improve productivity and ensure the benefits of economic growth are shared across communities.

```{r}
#| label: RI_4A1_no_qualifications_plot
#| eval: true
#| code-summary: "RI_4A1_no_qualifications line chart"
#| fig-alt: |
#|   Line chart of the percentage of residents aged 16 to 64 with no
#|   qualifications, from December 2022 onwards, for Bath and North East
#|   Somerset, Bristol, North Somerset, South Gloucestershire and the
#|   West of England as a whole. Lines are broadly flat to declining across
#|   the local authorities, with the West of England tracking below most
#|   of the individual unitary authorities.

RI_4A1_no_qualifications_plot
```
## Residents with Level 4+ Qualifications

```{r}
#| label: RI_4A2_level4_plus_gt
#| eval: true
#| code-summary: "RI_4A2_level4_plus summary"

indicator_fact_tbl |>
  format_indicator_summary(
    "RI_4A2_level4_plus",
    dim_tbl = core_dim_data_tbl,
    title = "Residents with Level 4+ Qualifications",
    subtitle = "Working-age population qualified to Level 4 and above",
    polarity = indicator_polarity[["RI_4A2_level4_plus"]]
  )
```

### Overview

Higher-level qualifications are an important indicator of workforce capability and economic competitiveness. This indicator measures the proportion of working-age residents qualified to Level 4 and above. (Level 3 qualifications include A levels and equivalent qualifications, this indicator captures qualifications above this level). These include a range of higher education qualifications, from Higher National Certificates and Certificates of Higher Education through to bachelor's degrees (Level 6) and postgraduate qualifications. Higher levels of attainment support innovation, productivity and growth in knowledge intensive sectors. The indicator reflects the region's ability to develop and attract the talent needed to support future economic growth.

```{r}
#| label: RI_4A2_level4_plus_plot
#| eval: true
#| code-summary: "RI_4A2_level4_plus line chart"
#| fig-alt: |
#|   Line chart of the percentage of residents aged 16 to 64 qualified to
#|   Level 4 or above, from December 2022 onwards, for Bath and North East
#|   Somerset, Bristol, North Somerset, South Gloucestershire and the
#|   West of England as a whole. Bath and North East Somerset consistently
#|   shows the highest share of higher-level qualifications, while the
#|   West of England sits in the middle of the group.

RI_4A2_level4_plus_plot
```

## NEET 16-17

```{r}
#| label: RI_4B3_neet_16_17_gt
#| eval: true
#| code-summary: "RI_4B3_neet_16_17 summary"

indicator_fact_tbl |>
  format_indicator_summary(
    "RI_4B3_neet_16_17",
    dim_tbl = core_dim_data_tbl,
    title = "NEET 16-17",
    subtitle = "16-17 year olds not in education, employment or training",
    polarity = indicator_polarity[["RI_4B3_neet_16_17"]]
  )
```

### Overview

Participation in education, employment or training during the transition from school is important for long-term outcomes. This indicator measures the proportion of 16-17 year olds who are not in education, employment or training (NEET). Lower rates indicate that more young people are progressing into positive destinations and developing the skills needed for future careers. Reducing NEET levels helps strengthen the future workforce and improve social mobility.

```{r}
#| label: RI_4B3_neet_16_17_plot
#| eval: true
#| code-summary: "RI_4B3_neet_16_17 line chart"
#| fig-alt: |
#|   Line chart of the percentage of 16 to 17 year olds not in education,
#|   employment or training, by year from 2015 onwards, for Bath and North
#|   East Somerset, Bristol, North Somerset, South Gloucestershire and the
#|   West of England as a whole. Rates vary between the local authorities
#|   year to year, with no single authority consistently highest or lowest.

RI_4B3_neet_16_17_plot
```
## Below Living Wage

```{r}
#| label: RI_4B5_below_living_wage_gt
#| eval: true
#| code-summary: "RI_4B5_below_living_wage summary"

indicator_fact_tbl |>
  format_indicator_summary(
    "RI_4B5_below_living_wage",
    dim_tbl = core_dim_data_tbl,
    title = "Below Living Wage",
    subtitle = "Proportion of employees earning below the Living Wage",
    polarity = indicator_polarity[["RI_4B5_below_living_wage"]]
  )
```

### Overview

The quality of employment is an important aspect of economic wellbeing. This indicator measures the proportion of employees earning below the Real Living Wage, as defined by the Living Wage Foundation. Lower levels indicate that a greater share of workers are earning wages that better reflect the cost of living. Monitoring this indicator helps assess whether economic growth is translating into improved living standards and the creation of quality employment opportunities. The most recent year of data is provisional and should be interpreted with caution, as revised estimates will be published the following year.

```{r}
#| label: RI_4B5_below_living_wage_plot
#| eval: true
#| code-summary: "RI_4B5_below_living_wage line chart"
#| fig-alt: |
#|   Bar chart of the percentage of employees in the West of England
#|   earning below the Real Living Wage, by year from 2015 onwards. The
#|   share below the Living Wage has generally declined over the period,
#|   though the most recent year is provisional and may be revised.

RI_4B5_below_living_wage_plot
```
## Economic Inactivity

```{r}
#| label: RI_4B2_economic_inactivity_gt
#| eval: true
#| code-summary: "RI_4B2_economic_inactivity summary"

indicator_fact_tbl |>
  format_indicator_summary(
    "RI_4B2_economic_inactivity",
    dim_tbl = core_dim_data_tbl,
    title = "Economic Inactivity",
    subtitle = "Proportion of working-age residents economically inactive",
    polarity = indicator_polarity[["RI_4B2_economic_inactivity"]]
  )
```
### Overview

Economic inactivity can limit workforce participation and reduce the availability of skills within the labour market. This indicator measures the proportion of working-age residents who are neither employed or actively seeking work. Lower levels of inactivity indicate greater participation in economic activity and access to employment opportunities. Reducing inactivity is an important part of addressing labour shortages and supporting inclusive growth.

```{r}
#| label: RI_4B2_economic_inactivity_plot
#| eval: true
#| code-summary: "RI_4B2_economic_inactivity line chart"
#| fig-alt: |
#|   Line chart of the economic inactivity rate for residents aged 16 to
#|   64, tracked monthly over the most recent years, comparing the
#|   West of England against Great Britain. The West of England rate
#|   tracks below the Great Britain rate for most of the period, with
#|   both series showing similar fluctuations over time.

RI_4B2_economic_inactivity_plot
```

## Apprenticeship Starts

```{r}
#| label: RI_4A3_apprenticeship_starts_gt
#| eval: true
#| code-summary: "RI_4A3_apprenticeship_starts annual change summary"

indicator_fact_tbl |>
  format_indicator_summary(
    "RI_4A3_apprenticeship_starts",
    dim_tbl = core_dim_data_tbl,
    title = "Annual change in apprenticeship starts",
    subtitle = "Percentage change from the previous academic year",
    polarity = indicator_polarity[["RI_4A3_apprenticeship_starts"]]
  )
```

### Overview

Apprenticeships provide an important route into skilled employment and career progression. This indicator measures the annual percentage change in apprenticeship starts across the region. Monitoring year-on-year change helps identify whether participation is increasing or decreasing over time and provides an indication of trends in apprenticeship uptake.

```{r}
#| label: RI_4A3_apprenticeship_starts_change_plot
#| eval: true
#| code-summary: "RI_4A3_apprenticeship_starts annual change line chart"
#| fig-alt: |
#|   Line chart of the year-on-year percentage change in apprenticeship
#|   starts for the West of England, from the 2016/17 academic year
#|   onwards, with a reference line at zero. The change fluctuates above
#|   and below zero across the years, showing no sustained upward or
#|   downward trend in apprenticeship uptake.

RI_4A3_apprenticeship_starts_change_plot
```

```{r}
#| label: RI_4A3_apprenticeship_starts_plot
#| eval: true
#| code-summary: "RI_4A3_apprenticeship_starts line chart"
#| fig-alt: |
#|   Stacked bar chart of apprenticeship starts by academic year from
#|   2015/16 onwards, with bars split by Bath and North East Somerset,
#|   Bristol, North Somerset and South Gloucestershire. Total starts vary
#|   year to year, with Bristol consistently contributing the largest
#|   share of the regional total.

RI_4A3_apprenticeship_starts_plot
```
## KS4 Attainment 8

```{r}
#| label: RI_4A4_ks4_attainment8_gt
#| eval: true
#| code-summary: "RI_4A4_ks4_attainment8 summary"

indicator_fact_tbl |>
  format_indicator_summary(
    "RI_4A4_ks4_attainment8",
    dim_tbl = core_dim_data_tbl,
    title = "KS4 Attainment 8",
    subtitle = "Average Attainment 8 score at Key Stage 4",
    polarity = indicator_polarity[["RI_4A4_ks4_attainment8"]]
  )
```

### Overview

Educational attainment provides the foundation for future learning, skills development and employment opportunities. This indicator measures [average pupil attainment across a range of subjects at the end of Key Stage 4 (GCSE)](https://www.gov.uk/government/publications/progress-8-school-performance-measure/secondary-accountability-measures-2025-guidance-for-maintained-secondary-schools-academies-and-free-schools#annex-a-progress-8-methodology). Strong educational outcomes help ensure young people are equipped to progress into further education, training and employment. Comparisons over time should be interpreted with some caution, as educational attainment and assessment arrangements were affected by the COVID-19 pandemic, particularly during 2020 and 2021.

```{r}
#| label: RI_4A4_ks4_attainment8_plot
#| eval: true
#| code-summary: "RI_4A4_ks4_attainment8 line chart"
#| fig-alt: |
#|   Line chart of average Attainment 8 score by academic year from
#|   2015/16 onwards, for Bath and North East Somerset, Bristol, North
#|   Somerset, South Gloucestershire and the West of England as a whole.
#|   Scores dip in the years affected by the COVID-19 pandemic before
#|   partially recovering, with Bath and North East Somerset generally
#|   scoring highest and the West of England in the middle of the group.

RI_4A4_ks4_attainment8_plot
```

## NEET 18-24
Please note: This indicator is currently under development and data is not yet available for reporting. Analysis will be included in a future update of the Regional Indicators report.

## Claimant Count in Deprived Wards
Please note: This indicator is currently under development and data is not yet available for reporting. Analysis will be included in a future update of the Regional Indicators report.

## Skills Gaps

Please note: This indicator is currently under development and data is not yet available for reporting. Analysis will be included in a future update of the Regional Indicators report.

