---
title: "Affordable and Sustainable Homes"
subtitle: "Building affordable and attractive homes in sustainable communities"
author: "CJ"
date: last-modified
execute:
  echo: true
  warning: false
  message: false
---

## Context

This chapter tracks indicators related to housing affordability and sustainability across the West of England. Indicators cover metrics like housing completions, affordability ratios, energy efficiency, social housing supply, and planning permissions. Indicators were developed to provide high-level insights into regional trends and performance in relation to the West of England's housing priorities as outlined in our [Growth Strategy](https://www.westofengland-ca.gov.uk/wp-content/uploads/2025/09/WE4915-Growth-Strategy-exec-summary_aw_web_v1.pdf).

```{r}
#| label: setup
#| code-summary: "Load required packages and source files"
#| include: false

# Load required packages
library(tidyverse)

# Source shared helpers (theme, helpers, fact_helpers, collate_fact)
source(here::here("scripts", "R", "_common.R"))
source(here::here("scripts", "R", "03-place", "RI_3A2_homes_epc_c_plus.R"))

indicator_fact_tbl <- collate_fact() |>
  build_reporting_view()

indicator_polarity <- core_dim_data_tbl |>
  dplyr::select(indicator_id, polarity) |>
  tibble::deframe()
```
---

## Energy Performance of Homes

```{r}
#| label: RI_3A2_homes_epc_c_plus_gt
#| eval: true
#| code-summary: "RI_3A2_homes_epc_c_plus summary"

indicator_fact_tbl |>
  format_indicator_summary(
    "RI_3A2_homes_epc_c_plus",
    title = "Energy Performance of Homes",
    subtitle = "Proportion of homes with EPC rating of C or above",
    units = "%",
    polarity = indicator_polarity[["RI_3A2_homes_epc_c_plus"]]
  )
```

### Overview

This indicator tracks the proportion of homes in the West of England with an Energy Performance Certificate (EPC) rating of C or above. A higher proportion indicates better energy efficiency and sustainability in the housing stock. Only about 70% of homes have an EPC, so this indicator is based on the subset of homes with an EPC, which may not be fully representative of all homes in the region. This dataset is from [MHCLG](https://get-energy-performance-data.communities.gov.uk/) and is based on the recently revised scheme for EPC.

```{r}
#| label: RI_3A2_homes_epc_c_plus_plot
#| eval: true
#| code-summary: "RI_3A2_homes_epc_c_plus plot"
#| fig-alt: |
#|   Line chart showing the cumulative monthly proportion of homes in the
#|   West of England with an EPC energy rating of C or better, from 2016 to
#|   2026. The y-axis shows percentage from 0% to 100%; the x-axis shows
#|   date. The line rises steadily from around 36% in 2016 to approximately
#|   53% in 2026, showing consistent improvement but indicating that nearly
#|   half of assessed homes still fall below a C rating.

RI_3A2_plot
```

---
<!-- ## Indicator -->

<!-- ```{r} -->
<!-- #| label: _gt -->
<!-- #| eval: true -->
<!-- #| code-summary: "summary" -->

<!-- ``` -->

<!-- ### Overview -->

<!-- ```{r} -->
<!-- #| label: _plot -->
<!-- #| eval: true -->
<!-- #| code-summary: " plot" -->


<!-- ``` -->

<!-- ### Findings -->

<!-- --- -->
<!-- ## Indicator -->

<!-- ```{r} -->
<!-- #| label: _gt -->
<!-- #| eval: true -->
<!-- #| code-summary: "summary" -->

<!-- ``` -->

<!-- ### Overview -->

<!-- ```{r} -->
<!-- #| label: _plot -->
<!-- #| eval: true -->
<!-- #| code-summary: " plot" -->


<!-- ``` -->

<!-- ### Findings -->

<!-- --- -->
<!-- ## Indicator -->

<!-- ```{r} -->
<!-- #| label: _gt -->
<!-- #| eval: true -->
<!-- #| code-summary: "summary" -->

<!-- ``` -->

<!-- ### Overview -->

<!-- ```{r} -->
<!-- #| label: _plot -->
<!-- #| eval: true -->
<!-- #| code-summary: " plot" -->


<!-- ``` -->

<!-- ### Findings -->

<!-- --- -->
<!-- ## Indicator -->

<!-- ```{r} -->
<!-- #| label: _gt -->
<!-- #| eval: true -->
<!-- #| code-summary: "summary" -->

<!-- ``` -->

<!-- ### Overview -->

<!-- ```{r} -->
<!-- #| label: _plot -->
<!-- #| eval: true -->
<!-- #| code-summary: " plot" -->


<!-- ``` -->

<!-- ### Findings -->

<!-- --- -->
<!-- ## Indicator -->

<!-- ```{r} -->
<!-- #| label: _gt -->
<!-- #| eval: true -->
<!-- #| code-summary: "summary" -->

<!-- ``` -->

<!-- ### Overview -->

<!-- ```{r} -->
<!-- #| label: _plot -->
<!-- #| eval: true -->
<!-- #| code-summary: " plot" -->


<!-- ``` -->

<!-- ### Findings -->

<!-- --- -->
<!-- ## Indicator -->

<!-- ```{r} -->
<!-- #| label: _gt -->
<!-- #| eval: true -->
<!-- #| code-summary: "summary" -->

<!-- ``` -->

<!-- ### Overview -->

<!-- ```{r} -->
<!-- #| label: _plot -->
<!-- #| eval: true -->
<!-- #| code-summary: " plot" -->


<!-- ``` -->

<!-- ### Findings -->

<!-- --- -->
<!-- ## Indicator -->

<!-- ```{r} -->
<!-- #| label: _gt -->
<!-- #| eval: true -->
<!-- #| code-summary: "summary" -->

<!-- ``` -->

<!-- ### Overview -->

<!-- ```{r} -->
<!-- #| label: _plot -->
<!-- #| eval: true -->
<!-- #| code-summary: " plot" -->


<!-- ``` -->

<!-- ### Findings -->

<!-- --- -->
<!-- ## Indicator -->

<!-- ```{r} -->
<!-- #| label: _gt -->
<!-- #| eval: true -->
<!-- #| code-summary: "summary" -->

<!-- ``` -->

<!-- ### Overview -->

<!-- ```{r} -->
<!-- #| label: _plot -->
<!-- #| eval: true -->
<!-- #| code-summary: " plot" -->


<!-- ``` -->

<!-- ### Findings -->

<!-- --- -->
<!-- ## Indicator -->

<!-- ```{r} -->
<!-- #| label: _gt -->
<!-- #| eval: true -->
<!-- #| code-summary: "summary" -->

<!-- ``` -->

<!-- ### Overview -->

<!-- ```{r} -->
<!-- #| label: _plot -->
<!-- #| eval: true -->
<!-- #| code-summary: " plot" -->


<!-- ``` -->

<!-- ### Findings -->

<!-- --- -->
<!-- ## Indicator -->

<!-- ```{r} -->
<!-- #| label: _gt -->
<!-- #| eval: true -->
<!-- #| code-summary: "summary" -->

<!-- ``` -->

<!-- ### Overview -->

<!-- ```{r} -->
<!-- #| label: _plot -->
<!-- #| eval: true -->
<!-- #| code-summary: " plot" -->


<!-- ``` -->

<!-- ### Findings -->

<!-- --- -->


## Summary

[Summary of key findings across all housing indicators. What is the overall picture for affordable and sustainable homes?]
