---
title: "Public Transport Connectivity"
subtitle: "Connecting the region through public transport & active travel"
author: "HB"
date: last-modified
execute:
  echo: true
  warning: false
  message: false
---

## Context

This chapter tracks indicators related to public transport connectivity across the West of England. The chapter draws on a range of national and local data sources to provide a high-level overview of regional trends and performance in relation to transport connectivity. Indicators were defined as part of the development of our [Growth Strategy](https://www.westofengland-ca.gov.uk/wp-content/uploads/2025/09/WE4915-Growth-Strategy-exec-summary_aw_web_v1.pdf).

## Setup

```{r}
#| label: setup
#| code-summary: "Load required packages"

# Load required packages
library(tidyverse)

# Source shared helpers (theme, helpers, fact_helpers, collate_fact)
source(here::here("scripts", "R", "_common.R"))
```

---

## Indicator 1: [Your Transport Indicator]

**Data source:** [Specify source]
**Last updated:** [Date]
**Refresh schedule:** [Annual/Quarterly/Monthly]


### Overview

[Describe what this indicator measures and why it matters for public transport connectivity]

```{r}
#| label: indicator-1
#| eval: false
#| code-summary: "Indicator 1 analysis"

# Your code here
# See chapters/01-economy/index.qmd for a worked example pattern:
# 1. Load data using load_csv()
# 2. Process/filter/aggregate data
# 3. Create visualization with theme_weca()
# 4. Calculate key statistics using format_number()
```

### Findings

[Your narrative findings here]

---

## Indicator 2: [Your Transport Indicator]

**Data source:**
**Last updated:**
**Refresh schedule:**

[Repeat structure - add up to 10 indicators total]

---

## Summary

[Summary of key findings across all transport indicators. What is the overall picture for public transport connectivity?]
